### 2017-03-20 paper/sections/Overview.tex
Tex.

\section{Overview}

\begin{itemize}
	\item Code for the problem (Draw example)
	\item Potential fixes (renaming, traits, ....)
	\item How it is solved with our language (client code, simple explanation) + show update
\end{itemize}