### 2017-03-20 paper/sections/RelatedWork.tex

\section{Related Work}

\begin{itemize}
	\item Multiple inheritance models
		\begin{itemize}
			\item Mixin
			\item Scala Mixins
			\item trait
			\item C++ model
			\item Java 8
			\item CZ
		\end{itemize}
	\item Resolve unintended method confliction
		\begin{itemize}
			\item Parents are Shared Parts of Objects: Inheritance and Encapsulation in SELF* (tiebreaker rule)
			\item ???
		\end{itemize}
	\item Static+Dynamic type method lookup (any existing language that supports this?)
	\item Formalization based on FJ (novelty: keep static types <I> in formalization)
		\begin{itemize}
			\item Existing formalizations based on FJ proposed new features and added rules in syntax and semantics. But we not only add rules, but also piggyback static types on almost all semantic rules to model method lookup. 
			\item Featherweight defenders, ...
		\end{itemize}
\end{itemize}