### 2017-03-28 paper/sections/Overview.tex
changes..

\section{Overview}

\begin{itemize}
	\item Code for the problem (Draw example)
	\item Potential fixes (renaming, traits, ....)
	\item How it is solved with our language (client code, simple explanation) + show update
\end{itemize}

\begin{lstlisting}
interface Deck {
   draw();  // draws a card from the Deck
}

interface Drawable {
   draw();  // draws something on the screen
}

interface DrawableDeck extends Deck, Drawable {
   ...
} 
\end{lstlisting}

\subsection{Potential fixes}

\begin{itemize}
    \item Explicit override m() in C.
    \item Choose one (such as Mixin model).
    \item With Traits: method exclusion.
    \item With Traits: method rename.
\end{itemize}
In previous solutions, only renaming can preserve the two different methods m(), 
but it’s cumbersome to do in practice: new names have to be introduced, and (probably) need to prefer one over another. 
Other solutions only preserve one method. Can we keep both methods without renaming?

\subsection{Solution in our language}
\begin{lstlisting}

\end{lstlisting}

\subsection{Method refinement}