### 2017-05-19 SpringerLNCS/paper.tex
change format to Springer LNCS

\documentclass{llncs}

\usepackage{listings}

\usepackage{xspace}
\usepackage{multicol}
\usepackage{microtype}%if unwanted, comment out or use option "draft"
%\usepackage[table,xcdraw]{xcolor} -- clash with preamble.tex
\usepackage{color}
% \usepackage{amsthm}
% \usepackage{amsmath}
\usepackage{stmaryrd}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{fancyvrb}
\usepackage{url}
%\usepackage{pstricks,pst-node,pst-tree} -- clash with preamble.tex
\usepackage{bbm}
%\usepackage{pgf} -- clash with preamble.tex
\usepackage{multirow}
\usepackage{enumitem}

\usepackage{verbatim}
\usepackage{graphicx}
\usepackage{wrapfig}
\usepackage[normalem]{ulem}

\usepackage[T1]{fontenc}
\usepackage[scaled=0.85]{beramono}
% \usepackage{mathpartir}
\usepackage[utf8]{inputenc}
\usepackage{flushend}

\let\proof\relax
\let\endproof\relax
\let\mathpar\relax

\input{predef/names.tex}
\input{predef/mim/syntax.tex}
\input{predef/others/preamble.tex}
\input{predef/macros/common_pl_macros.tex}
\input{predef/algebra/syntax.tex}
\input{predef/f-and/base.tex}
\input{predef/f-and/orthogonality.tex}
\input{predef/f-and/wellformedness.tex}
\input{predef/f-and/subtyping.tex}
\input{predef/f-and/select.tex}
\input{predef/f-and/restrict.tex}
\input{predef/f-and/typing.tex}

\begin{document}

\title{NoTitle}
%\subtitle{\InterfaceBased Programming for the Masses}

% \authorinfo{NoAuthor}
%            {The University of Hong Kong, China}
% %           {\{ylwang,hyzhang,bruno\}@cs.hku.hk}
% %\authorinfo{Marco Servetto}
% %          {Victoria University of Wellington, New Zealand}
% %          {marco.servetto@ecs.vuw.ac.nz}
\author{Author 1, Author 2}

\institute{Lab, University, Address}

\maketitle

\begin{abstract}
  Text.
\end{abstract}

\input{sections/Introduction.tex}

\input{sections/Overview.tex}

\input{sections/Formalization.tex}

\input{sections/Implementation.tex}

\input{sections/Discussion.tex}

\input{sections/CaseStudy.tex}

\input{sections/RelatedWork.tex}

\input{sections/Conclusion.tex}

\section*{Acknowledgments}\label{sec:Acknowledgments}

Authors would like to thank YYYYY.

% \newpage
% \clearpage
% \bibliographystyle{abbrvnat}
\bibliographystyle{splncs03}
\bibliography{paper}

% \appendix
% \input{sections/Appendix_Formalization.tex}
% \input{sections/Translation.tex}
% \input{sections/Appendix_Translation.tex}
% \input{sections/Appendix.tex}

\end{document}