### 2017-05-19 SpringerLNCS/sections/Introduction.tex

\section{Introduction}

\begin{itemize}
	\item problem: Unintended method confliction in Multiple inheritance
	\item Existing approaches or models and Their drawbacks
	\item New features (update)
	\item Our contributions
\end{itemize}

In multiple inheritance, naming conflicts often occurs. Among these conflicts, some are real conflicts which needs 
explicit resolve by programmers, however, there are cases where accidental naming conflicts occurs, where the conflicting
methods have completely different meaning/domain which just share the same name.

Existing OOP models have taken care of the first case intensively. However, few of them supports 
unintended method confliction well. Trait and other 
mainstream OO models do not allow unintended methods confliction to co-exist. 
SELF~\cite{} uses \emph{sender path tiebreaker rule} to automatically resolve 
ambiguities that are almost certainly caused by accidental naming conflicts. C++ allows methods with the same signature 
co-exist in a class via inheritance and programmers can use $::$ operator to select the method wanted. 
However none of them allows refining these unintended conflicting methods in subclasses. We propose a calculus that 
deals with unintended method confliction and meanwhile allows refining these methods.

Contributions:
\begin{itemize}
    \item A multiple inheritance model formalized as \MIM.
    \item Novel notion \updates for method path updating.
    \item Implementation of a simple typechecker and evaluator in Scala.
\end{itemize}