### 2017-05-21 SpringerLNCS/paper.tex
Overview.

\documentclass{llncs}

\usepackage{listings}

\usepackage{xspace}
\usepackage{multicol}
\usepackage{microtype}%if unwanted, comment out or use option "draft"
%\usepackage[table,xcdraw]{xcolor} -- clash with preamble.tex
\usepackage{color}
% \usepackage{amsthm}
% \usepackage{amsmath}
\usepackage{stmaryrd}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{fancyvrb}
\usepackage{url}
%\usepackage{pstricks,pst-node,pst-tree} -- clash with preamble.tex
\usepackage{bbm}
%\usepackage{pgf} -- clash with preamble.tex
\usepackage{multirow}
\usepackage{enumitem}

\usepackage{verbatim}
\usepackage{graphicx}
\usepackage{wrapfig}
\usepackage[normalem]{ulem}

\usepackage[T1]{fontenc}
\usepackage[scaled=0.85]{beramono}
% \usepackage{mathpartir}
\usepackage[utf8]{inputenc}
\usepackage{flushend}
\usepackage{lmodern}

\let\proof\relax
\let\endproof\relax
\let\mathpar\relax

\input{predef/names.tex}
\input{predef/mim/syntax.tex}
\input{predef/others/preamble.tex}
\input{predef/macros/common_pl_macros.tex}
\input{predef/f-and/base.tex}
\input{predef/f-and/orthogonality.tex}
\input{predef/f-and/wellformedness.tex}
\input{predef/f-and/subtyping.tex}
\input{predef/f-and/select.tex}
\input{predef/f-and/restrict.tex}
\input{predef/f-and/typing.tex}

\lstset{ %
	language=Java,                % choose the language of the code
	columns=flexible,
	lineskip=-1pt,
	basicstyle=\ttfamily\small,       % the size of the fonts that are used for the code
	numbers=none,                   % where to put the line-numbers
	numberstyle=\ttfamily\tiny,      % the size of the fonts that are used for the line-numbers
	stepnumber=1,                   % the step between two line-numbers. If it's 1 each line will be numbered
	numbersep=5pt,                  % how far the line-numbers are from the code
	backgroundcolor=\color{white},  % choose the background color. You must add \usepackage{color}
	showspaces=false,               % show spaces adding particular underscores
	showstringspaces=false,         % underline spaces within strings
	showtabs=false,                 % show tabs within strings adding particular underscores
	morekeywords={var},
	%  frame=single,                   % adds a frame around the code
	tabsize=2,                  % sets default tabsize to 2 spaces
	captionpos=none,                   % sets the caption-position to bottom
	breaklines=true,                % sets automatic line breaking
	breakatwhitespace=false,        % sets if automatic breaks should only happen at whitespace
	title=\lstname,                 % show the filename of files included with \lstinputlisting; also try caption instead of title
	escapeinside={(*}{*)},          % if you want to add a comment within your code
	keywordstyle=\ttfamily\bfseries,
	aboveskip=0pt,
	belowskip=0pt
	% commentstyle=\color{Gray},
	% stringstyle=\color{Green}
}

\begin{document}

\title{NoTitle}
%\subtitle{\InterfaceBased Programming for the Masses}

% \authorinfo{NoAuthor}
%            {The University of Hong Kong, China}
% %           {\{ylwang,hyzhang,bruno\}@cs.hku.hk}
% %\authorinfo{Marco Servetto}
% %          {Victoria University of Wellington, New Zealand}
% %          {marco.servetto@ecs.vuw.ac.nz}
\author{Author 1, Author 2}

\institute{Lab, University, Address}

\maketitle

\begin{abstract}
  Text.
\end{abstract}

\input{sections/Introduction.tex}

\input{sections/Overview.tex}

\input{sections/Formalization.tex}

\input{sections/Implementation.tex}

\input{sections/Discussion.tex}

\input{sections/CaseStudy.tex}

\input{sections/RelatedWork.tex}

\input{sections/Conclusion.tex}

\section*{Acknowledgments}\label{sec:Acknowledgments}

Authors would like to thank YYYYY.

% \newpage
% \clearpage
% \bibliographystyle{abbrvnat}
\bibliographystyle{splncs03}
\bibliography{paper}

% \appendix
% \input{sections/Appendix_Formalization.tex}
% \input{sections/Translation.tex}
% \input{sections/Appendix_Translation.tex}
% \input{sections/Appendix.tex}

\end{document}