### 2017-06-05 SpringerLNCS/sections/Conclusion.tex
Conclusion.

\section{Conclusion}

This paper proposes \MIM{} as a new multiple inheritance model for unintentional method conflicts.
Existing approaches including static dispatch and dynamic dispatch have to compromise between code reuse
and type safety, whereas \MIM{} uses a different solution called hierarchical dispatch to obtain both.
Such an approach not only offers great code reuse like dynamic dispatch, but also ensures unambiguity by
our algorithm for method binding. This paper also introduces method updates for refinements on branches individually.
\MIM{} is formalized with a basis on Featherweight Java, and supported
by a few theorems.

Our model can certainly be improved at some aspects. We did not formalize fields, though it is difficult to
put state and multiple inheritance together []. For simplicity, we did not formalize other common features including
method overloading, casts, covariant return types, and so on. Moreover, we restrict method updates to target only at the
original methods. Potentially a looser condition can better support encapsulation and modularity with respect to code design.