### 2017-09-06 NewVersion/paper.tex
Marco first commit for formalization

\documentclass{llncs}

\usepackage{listings}

\usepackage{xspace}
\usepackage{multicol}
\usepackage{microtype}%if unwanted, comment out or use option "draft"
%\usepackage[table,xcdraw]{xcolor} -- clash with preamble.tex
\usepackage{color}
% \usepackage{amsthm}
% \usepackage{amsmath}
\usepackage{stmaryrd}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{fancyvrb}
\usepackage{url}
%\usepackage{pstricks,pst-node,pst-tree} -- clash with preamble.tex
\usepackage{bbm}
%\usepackage{pgf} -- clash with preamble.tex
\usepackage{multirow}
\usepackage{enumitem}

\usepackage{verbatim}
\usepackage{graphicx}
\usepackage{wrapfig}
\usepackage[normalem]{ulem}

\usepackage[T1]{fontenc}
\usepackage[scaled=0.85]{beramono}
% \usepackage{mathpartir}
\usepackage[utf8]{inputenc}
\usepackage{flushend}
\usepackage{lmodern}

\let\proof\relax
\let\endproof\relax
\let\mathpar\relax

%%\newcommand\bruno[1]{\authornote{bruno}{red}{#1}}

\input{predef/names.tex}
\input{predef/mim/syntax.tex}
\input{predef/others/preamble.tex}
\input{predef/macros/common_pl_macros.tex}
\input{predef/f-and/base.tex}
\input{predef/f-and/orthogonality.tex}
\input{predef/f-and/wellformedness.tex}
\input{predef/f-and/subtyping.tex}
\input{predef/f-and/select.tex}
\input{predef/f-and/restrict.tex}
\input{predef/f-and/typing.tex}

\newcommand{\authornote}[3]{{\color{#2} {\sc #1}: #3}}
\newcommand{\authorText}[2]{{\color{#1}#2}}

 \newcommand\bruno[1]{\authornote{bruno}{red}{#1}}
 \newcommand\yanlin[1]{\authornote{yanlin}{purple}{#1}}
 \newcommand\marco[1]{\authornote{marco}{blue}{#1}}
 \newcommand\marcoT[1]{\authorText{blue}{#1}}
 \newcommand\haoyuan[1]{\authornote{haoyuan}{orange}{#1}}

%\newcommand\bruno[1]{}
%\newcommand\yanlin[1]{}
%\newcommand\marco[1]{}
%\newcommand\haoyuan[1]{}

\newcommand\saveSpaceFig{\vspace{-2ex}}

\lstset{ %
	language=Java,                % choose the language of the code
	columns=flexible,
	lineskip=-1pt,
	basicstyle=\ttfamily\small,       % the size of the fonts that are used for the code
	numbers=none,                   % where to put the line-numbers
	numberstyle=\ttfamily\tiny,      % the size of the fonts that are used for the line-numbers
	stepnumber=1,                   % the step between two line-numbers. If it's 1 each line will be numbered
	numbersep=5pt,                  % how far the line-numbers are from the code
	backgroundcolor=\color{white},  % choose the background color. You must add \usepackage{color}
	showspaces=false,               % show spaces adding particular underscores
	showstringspaces=false,         % underline spaces within strings
	showtabs=false,                 % show tabs within strings adding particular underscores
	morekeywords={var,updates},
	%  frame=single,                   % adds a frame around the code
	tabsize=2,                  % sets default tabsize to 2 spaces
	captionpos=none,                   % sets the caption-position to bottom
	breaklines=true,                % sets automatic line breaking
	breakatwhitespace=false,        % sets if automatic breaks should only happen at whitespace
	title=\lstname,                 % show the filename of files included with \lstinputlisting; also try caption instead of title
	escapeinside={(*}{*)},          % if you want to add a comment within your code
	keywordstyle=\ttfamily\bfseries,
	aboveskip=0pt,
	belowskip=0pt
	% commentstyle=\color{Gray},
	% stringstyle=\color{Green}
}

\begin{document}

% \title{Untangling Unintended Conflicts with Hierarchical Dispatch}
% \title{FHJ: A Formal Model for Resolving Unintended Conflicts in Multiple Inheritance}
\title{Hierarchical Dispatching and Overriding}
%\subtitle{\InterfaceBased Programming for the Masses}

% \authorinfo{NoAuthor}
%            {The University of Hong Kong, China}
% %           {\{ylwang,hyzhang,bruno\}@cs.hku.hk}
% %\authorinfo{Marco Servetto}
% %          {Victoria University of Wellington, New Zealand}
% %          {marco.servetto@ecs.vuw.ac.nz}
\author{Authors Omitted}

\institute{}

\maketitle

\vspace{-4ex}

\begin{abstract}
Multiple inheritance is a valuable feature for Object-Oriented
Programming. However, it is also tricky to get right, as illustrated by
the extensive literature on the topic. One of the most promising 
approaches to multiple inheritance is the \emph{trait} model. Traits offer a
restricted model of multiple inheritance that is easy to reason and 
have many elegant properties. Traits have good support for method 
conflicts which arise from \emph{diamond inheritance}:
conflicts that arise from method implementations sharing a common 
ancestor. However, the mechanisms of traits are inadequate to deal 
with \emph{unintentional method conflicts}: conflicts which 
arise from two unrelated methods that happen to share the same name
and signature.

This paper presents a new model \name{} (short for \emph{\textbf{F}eatherweight \textbf{H}ierarchical \textbf{J}ava}) as an extension/variation to the trait model that deals with unintentional method conflicts. The method resolution semantics resembles C++, so part of our formalization can be used as a partial C++ method lookup algorithm formalization. We discuss the similarities and differences with C++ and also advantages and disadvantages. In our new model, conflicting methods arising from unrelated methods can coexist in the same component. Moreover, one novel feature we add to our model is hierarchical method overriding: methods can be \emph{independently overridden} along multiple inheritance hierarchy. What ensures unambiguity is the use of information about the class hierarchy. In turn, hierarchical information is also employed in method dispatching, which uses a combination of static and dynamic type information to choose the implementation of a method at run-time. We give illustrative examples of our language and features and formalize \name{} as a minimal Featherweight-Java style calculus.

\end{abstract}

\input{sections/Introduction.tex}
\input{sections/Overview.tex}
\input{sections/Formalization.tex}
\input{sections/Auxiliary.tex}
\input{sections/Discussion.tex}
\input{sections/RelatedWork.tex}
\input{sections/Conclusion.tex}

\begin{comment}
\section*{Acknowledgments}\label{sec:Acknowledgments}

Authors would like to thank YYYYY.
\end{comment}

% \newpage
% \clearpage
% \bibliographystyle{abbrvnat}
\bibliographystyle{splncs03}
\bibliography{paper}

\newpage
\appendix
\input{sections/Appendix.tex}
% \input{sections/Appendix_Formalization.tex}
% \input{sections/Translation.tex}
% \input{sections/Appendix_Translation.tex}

\end{document}