### 2017-09-30 NewVersion/paper.tex
Abstract

\documentclass{llncs}

\usepackage{listings}

\usepackage{xspace}
\usepackage{multicol}
\usepackage{microtype}%if unwanted, comment out or use option "draft"
%\usepackage[table,xcdraw]{xcolor} -- clash with preamble.tex
\usepackage{color}
% \usepackage{amsthm}
% \usepackage{amsmath}
\usepackage{stmaryrd}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{fancyvrb}
\usepackage{url}
%\usepackage{pstricks,pst-node,pst-tree} -- clash with preamble.tex
\usepackage{bbm}
%\usepackage{pgf} -- clash with preamble.tex
\usepackage{multirow}
\usepackage{enumitem}

\usepackage{verbatim}
\usepackage{graphicx}
\usepackage{wrapfig}
\usepackage[normalem]{ulem}

\usepackage[T1]{fontenc}
\usepackage[scaled=0.85]{beramono}
% \usepackage{mathpartir}
\usepackage[utf8]{inputenc}
\usepackage{flushend}
\usepackage{lmodern}

\let\proof\relax
\let\endproof\relax
\let\mathpar\relax

%%\newcommand\bruno[1]{\authornote{bruno}{red}{#1}}

\input{predef/names.tex}
\input{predef/mim/syntax.tex}
\input{predef/others/preamble.tex}
\input{predef/macros/common_pl_macros.tex}
\input{predef/f-and/base.tex}
\input{predef/f-and/orthogonality.tex}
\input{predef/f-and/wellformedness.tex}
\input{predef/f-and/subtyping.tex}
\input{predef/f-and/select.tex}
\input{predef/f-and/restrict.tex}
\input{predef/f-and/typing.tex}

\newcommand{\authornote}[3]{{\color{#2} {\sc #1}: #3}}
\newcommand{\authorText}[2]{{\color{#1}#2}}

 \newcommand\bruno[1]{\authornote{bruno}{red}{#1}}
 \newcommand\yanlin[1]{\authornote{yanlin}{purple}{#1}}
 \newcommand\marco[1]{\authornote{marco}{blue}{#1}}
 \newcommand\marcoT[1]{\authorText{blue}{#1}}
 \newcommand\haoyuan[1]{\authornote{haoyuan}{orange}{#1}}

%\newcommand\bruno[1]{}
%\newcommand\yanlin[1]{}
%\newcommand\marco[1]{}
%\newcommand\haoyuan[1]{}

\newcommand\saveSpaceFig{\vspace{-2ex}}

\lstset{ %
	language=Java,                % choose the language of the code
	columns=flexible,
	lineskip=-1pt,
	basicstyle=\ttfamily\small,       % the size of the fonts that are used for the code
	numbers=none,                   % where to put the line-numbers
	numberstyle=\ttfamily\tiny,      % the size of the fonts that are used for the line-numbers
	stepnumber=1,                   % the step between two line-numbers. If it's 1 each line will be numbered
	numbersep=5pt,                  % how far the line-numbers are from the code
	backgroundcolor=\color{white},  % choose the background color. You must add \usepackage{color}
	showspaces=false,               % show spaces adding particular underscores
	showstringspaces=false,         % underline spaces within strings
	showtabs=false,                 % show tabs within strings adding particular underscores
	morekeywords={var,updates},
	%  frame=single,                   % adds a frame around the code
	tabsize=2,                  % sets default tabsize to 2 spaces
	captionpos=none,                   % sets the caption-position to bottom
	breaklines=true,                % sets automatic line breaking
	breakatwhitespace=false,        % sets if automatic breaks should only happen at whitespace
	title=\lstname,                 % show the filename of files included with \lstinputlisting; also try caption instead of title
	escapeinside={(*}{*)},          % if you want to add a comment within your code
	keywordstyle=\ttfamily\bfseries,
	aboveskip=0pt,
	belowskip=0pt
	% commentstyle=\color{Gray},
	% stringstyle=\color{Green}
}

\begin{document}

% \title{Untangling Unintended Conflicts with Hierarchical Dispatch}
% \title{FHJ: A Formal Model for Resolving Unintended Conflicts in Multiple Inheritance}
\title{Hierarchical Dispatching and Overriding}
%\subtitle{\InterfaceBased Programming for the Masses}

% \authorinfo{NoAuthor}
%            {The University of Hong Kong, China}
% %           {\{ylwang,hyzhang,bruno\}@cs.hku.hk}
% %\authorinfo{Marco Servetto}
% %          {Victoria University of Wellington, New Zealand}
% %          {marco.servetto@ecs.vuw.ac.nz}
\author{Authors Omitted}

\institute{}

\maketitle

\vspace{-4ex}

\begin{abstract}
Multiple inheritance is a valuable feature for Object-Oriented
Programming. However, it is also tricky to get right, as illustrated by
the extensive literature on the topic. A key issue 
is the \emph{ambiguity} arising from inheriting multiple parents,
which can have conflicting methods and fields. 
Numerous existing work provides solutions for 
conflicts which arise from \emph{diamond inheritance}: i.e.
conflicts that arise from implementations sharing a common 
ancestor. However, most mechanisms are inadequate to deal 
with \emph{unintentional method conflicts}: conflicts which 
arise from two unrelated methods that happen to share the same name
and signature.

\begin{comment}
One of the most promising 
approaches to multiple inheritance is the \emph{trait} model. Traits offer a
restricted model of multiple inheritance that is easy to reason and 
have many elegant properties. Traits have good support for method 
conflicts which arise from \emph{diamond inheritance}:
conflicts that arise from method implementations sharing a common 
ancestor. However, the mechanisms of traits are inadequate to deal 
with \emph{unintentional method conflicts}: conflicts which 
arise from two unrelated methods that happen to share the same name
and signature. 
\end{comment}

This paper presents a new model called \emph{\textbf{F}eatherweight
  \textbf{H}ierarchical \textbf{J}ava} (\name{}) that deals with
unintentional method conflicts.
In our
new model, conflicting methods arising from unrelated methods can
coexist in the same component. Furthermore our model supports
\emph{hierarchical method overriding}: that is, methods can be
\emph{independently overridden} along the multiple inheritance
hierarchy. To avoid ambiguity, hierarchical information is 
employed in method dispatching, which uses a combination of static and
dynamic type information to choose the implementation of a method at
run-time. We give illustrative examples of our language and features
and formalize \name{} as a minimal Featherweight-Java style calculus.

\begin{comment}
Furthermore we discuss similarities and differences to

What ensures unambiguity is the use of information about
the class hierarchy.

\name{} is partly inspired by the method resolution semantics
of C++, but it also incoorporates ideas from the trait model and Java
8's default methods. We discuss the similarities
and differences with and also advantages and disadvantages. 
\end{comment}

\end{abstract}

\input{sections/Introduction.tex}
\input{sections/Overview.tex}
\input{sections/Formalization.tex}
\input{sections/Auxiliary.tex}
\input{sections/Discussion.tex}
\input{sections/RelatedWork.tex}
\input{sections/Conclusion.tex}

\begin{comment}
\section*{Acknowledgments}\label{sec:Acknowledgments}

Authors would like to thank YYYYY.
\end{comment}

% \newpage
% \clearpage
% \bibliographystyle{abbrvnat}
\bibliographystyle{splncs03}
\bibliography{paper}

\newpage
\appendix
\input{sections/Appendix.tex}
% \input{sections/Appendix_Formalization.tex}
% \input{sections/Translation.tex}
% \input{sections/Appendix_Translation.tex}

\end{document}