### 2017-12-08 NewVersion/response.txt
separate response

We thank the reviewers for the valuable comments. We try to address the reviews one by one.

Reviewer A.
- generalize the method overriding semantics
We think this suggestion is useful and valueble, we didn't think about this before. It does produce a
simpler method overriding semantics and definitely can be a future work.

- "renaming can break the subtyping relation"
Naturally renaming can break the subtyping relation, of course, there might be techniques as you said 
that tracks the rename mapping and keeps the subtyping relation, but that's the special case and we 
didn't discuss it in the paper.

- "you do not mention the more direct way in which C++ allows disambiguating methods by qualification..."
In Section 2.1 we have mentioned the C++ qualified name approach, see Sec 2.1 -> Potential Fixes -> 
Static dispatch.

- "C++ allows both overriding and shadowing ..."
Since FHJ does not support overloading, it's model is much simpler than C++, so there's no concern 
of distinguishing overriding and shadowing of inherited methods. But we agree that it could be interesting
to add this feature.

- "the notion of 'original' method...."
Sorry that we didn't make it clear. The term "original method" is supposed to be relative to
"hierarchical overriding method". It does include overrides and maybe a better name (for example
"own method"?) is needed.

- "abusing override self syntax for non-(hierarchical-)override...can be empty?..."
Yes, `I { m() }` is a syntactic sugar for `I { m() override I ... }`. In a real language, this should be 
supported. But in formlization, for simplicity, we union them into one.

- Footnote 4: “new I()()” — macro fail?
For footnote 4, the version submitted is "In Java the corresponding syntax is new I(){}." Maybe the 
version you have is not updated?

- "Figure 6: The rule C-StaticType renders the reduction relation non-deterministic"
Without (C-FReduce), (C-StaticType) IS infinite. But in (C-FReduce), the condition `e \neq new J()`
prevent the infinite loop.

Reviewer B.
- "Eiffel seems to allow renaming ..."
In Eiffel, programmers still have to rename methods to resolve the conflicts, so I don't think it is a 
different approach and actually you can't keep the method with the SAME name.

> MiddleMan approach, "whether this problem is really worth a language extension..."
The example in "The C++ programming language" and stackoverflow are both similar to the MiddleMan
approach (they have the same spirit, just MiddleMan approach uses parameter types to distinguish
different methods, while the others use separate sets of methods to distinguish).
We agree that this name conflict problem is not quite popular in practice. That's also the reason
why we do not have a big case study in the paper, because we didn't find a senario that
unintentional method conflicts are frequently used in one project.   But once this issue is
encountered, the effect could be severe, and that existing workarounds might not help a lot. And
since the C++ books, the trait paper as well as other papers mentioned this issue, it might be a
sign that this problem is still an interesting issue in certain cases, which worth exploring and
taking into consideration.  Finally our work shows that this feature is lightweight, it does not
complicate the semantics a lot and can be easily added to language models.

- "It would strengthen your paper if you would provide an implementation..."
As mentioned in Section 1, we have a prototype implementation available here: 
https://github.com/YanlinWang/MIM/tree/master/Calculus

- "What has been removed from FJ: just downcasts?"
Also constructors and fields.

- "5.2: static invocation confusing..."
* "the method m in J_0 that hierarchically overrides J_1" means, the method m defined like this:
interface J_0 { K m() override J_1 { ... } }
* "e1.J_0@J_1 :: m(e2)" is different from "((J1) e1.m(e2)". The first is a static call while the
second dynamic. The result would be different when the method m gets overridden in an heir of J_0.
* For the question "whether it should be forbidden", we think it is a design choice. If the user
makes  a normal e.m() call, for sure it wouldn't access the super methods. But the static invocation
here gives users the freedom to directly call super methods.

- "p.22: "It breaks structural subtyping.."

Review C.
- "The proposal seems unnecessarily complex. In NextGen and MixGen ..."

- "what do you mean by 'reduction model similar to FJ'?"
Sorry for the English, what we mean is that the reduction steps follow the reduction rules in FJ-like
languages, where no static types are tracked.

- "you mention for the first time 'origin'..."
On page 8, "origin" means where the method is defined, we avoid using "original methods" because it
it not introduced yet on page 8. "original method" is a relative concept to hierarchical overriding 
methods.

- "Section 3.2: subtyping is ANTIsymmetric."
Yes, it is not symmetric, we will delete it.