### 2018-06-04 ECOOP18/sections/Discussion.tex
marco's suggestions

\section{Discussion}\label{sec:discussion}
In this section, we will discuss the design space and reflect about some of the design decisions of our work. We relate our language to traits, Java interfaces as well as other languages. Furthermore, we discuss ways to improve our work.

\yanlin{marco: The whole section 5 may benefit some English proofreading, but I'm too
	bad at English to really help... may be Bruno can give some more
	support here?}

\subsection{Abstract Methods}
Abstract methods are one of the key features in most general OO languages. For example, Java interfaces were designed
to include only method declarations, and those abstract methods can be implemented in a class body. 
The formal Featherweight Java model~\cite{Igarashi01FJ} does not include abstract methods because of the orthogonality
to the core calculus. In traits, a
similar idea is to use keywords like ``\textbf{require}'' for abstract method declarations~\cite{scharli03traits}.
Abstract methods provide a way to
delay the implementations to future subtypes. Using overriding, they also help to ``exclude'' existing implementations.

In our formalized calculus, however, abstract methods are not a completely orthogonal feature. The $\canInstantiate$ function
has to check whether an interface can be instantiated by looking at all the inherited branches and checking if each most specific method is concrete or not.

Our formalization has a simple form of abstract methods, which behave similarly to conventional methods with respect to conflicts.
Other languages may behave differently. 
For instance, in Java 8 when putting two identical abstract methods together by multiple inheritance, there is no conflict error. In Figure~\ref{fig:abstractdiamond}, we use italic $\emph{m}$ to denote abstract methods. In both cases, the Java compiler accepts the definition of $C$ and automatically merges the two inherited methods $m$ into a single one. \MIM{} behaves differently from Java in both cases.
In the \wordfork{} inheritance case (left), $C$ will have two distinct abstract methods corresponding to $A.m$ and $B.m$. 
In the diamond inheritance case, the definition of $C$ is rejected. 
There are two reasons for this difference in behaviour. Firstly, our formalization just treats abstract methods as concrete methods with an empty body, and that simplifies
the rules and proofs a lot. Secondly, and more importantly, we distinguish and treat differently conflicting methods, since they may represent different operations, even if they are abstract. Thus our model adopts a very conservative behavior rather than automatically merging 
methods by default (as done in many languages). 
\yanlin{717: we could mention that Godez (the guy maintaing java now) would
	prefer if Java behaved like our formalization but he can not for
	retrocompatibility.
	Re: where do you think is the proper place to mention this? }
Arguably, for the diamond case, as we have mentioned, it is actually an intentional conflict due to the same source $T$.
It is possible to change our model to account for other behaviors for abstract methods, but we view this as a mostly 
orthogonal change to our work, which should not affect the essence of the model presented here.

\begin{figure*}[t]
\saveSpaceFig
    % \nocaptionrule
    % \centering
    \centering
    \begin{minipage}[t]{0.45\textwidth}
        \centering
        \includegraphics[width=1.6cm]{pics/P7.pdf}
    \end{minipage}
    \centering
    \hspace*{2pt}
    \begin{minipage}[t]{0.45\textwidth}
        \centering
        \includegraphics[width=1.4cm]{pics/P8.pdf}
    \end{minipage}  
    \caption{\wordFork{} inheritance (left) and diamond inheritance (right) on abstract methods.}\label{fig:abstractdiamond}
\saveSpaceFig
\end{figure*}

\subsection{Orthogonal \& Non-Orthogonal Extensions}\label{sec:orthoext}
Our model is designed as a minimal calculus that focuses on resolving unintentional conflicts. Therefore, we have omitted a number of
common orthogonal features including primitive types, assignments, method overloading, covariant method return types, static dispatch, and so on.
Those features can, in principle, be modularly added to the model without breaking type soundness. For example, we present the additional syntax, typing and semantic rules of static invocation below as an extension:

\begin{mathpar}
    \begin{array}{llrl}
        \text{Expressions}  & e  & \Coloneqq & \ldots \; \mid \; e.J_0@J_1::m(\overline{e})
    \end{array} \\
    \tstaticinvk \\
    \sstaticinvk
\end{mathpar}
A static invocation $e.J_0@J_1::m(\overline{e})$ aims at finding the method $m$ in $J_0$ that hierarchically overrides $J_1$, thus $J_0[m\ \kwoverride\ J_1]$ is invoked. As shown in \textsc{(S-StaticInvk)}, static dispatch needs a receiver for the substitution of the ``$\kwthis$'' reference, so as to provide the latest implementations. In fact, static dispatch is common in OO programming, as it provides a shortcut to the reuse of old implementation easily, and super calls can also rely on this feature. For convenience we just make it simple above, whereas in languages like C++ or Java, the static or super invocations are more flexible, as they can climb the class hierarchy.

\yanlin{Bruno: discussed multiple overriding and the two designs}
One non-orthogonal extension to \MIM{} could be to generalize the model to allow multiple hierarchical method overrding, meaning that, we allow overriding methods to update multiple branches instead of only one branch. It offers a more fine-grained mechanism for merging, and can be helpful to easily understand the structure of the hierarchy. Multiple overriding would be useful in the following situation, for example:

\begin{lstlisting}
interface A { void m() {...} }
interface B { void m() {...} }
interface C { void m() {...} }
interface D extends A, B, C {
	void m() override A,B {...}   // overrides branches A and B only
	void m() override C {...}      // overrides branch C
}
\end{lstlisting}

Here \lstinline|D| inherits from three interfaces \lstinline|A, B, C| with
conflicting methods \lstinline|m|, but only merges two of those methods.
While we can simulate D without multiple overriding in our calculus
(by introducing an intermediate class), a better approach would be
to support multiple overriding natively.

We present the modification of syntax, typing and semantic rules below (abstract methods omitted):

\begin{mathpar}
	\begin{array}{llrl}
		    \text{Methods}      & M  & \Coloneqq & \ldots \;   \mid
       \momethod{I}{m}{I_x}{x}{J}{e}  
	\end{array} \\
	\tmomethod 
\end{mathpar}

Semantic rules themselves remain unchanged, however, we need to change slightly the definition of \mostSpecificOverride{} in \mbody:

\saveSpaceFig
\begin{flalign*}
& \rhd \textit{Definition of } \mostSpecificOverride(m, I, J): & \\
& \bullet \mostSpecificOverride(m, I, J) = \prune(overrides) & \\
& \indent\indent \textrm{with: } overrides = \{K \mid \subt{I}{K}, \; \subt{K}{J} \textrm{ and } K[m\ \kwoverride\ \overline{J}] \textrm{ where } J \in \overline{J} 
\end{flalign*}
\saveSpaceFig

With this approach, branches \lstinline|A| and \lstinline|B| are merged in the sense that they share the same code, which can be separately updated in future interfaces. Another approach would be to deeply merge the branches, with similar effect as introducing an intermediate interface \lstinline|AB| to explicitly merge the two branches. However, this approach is problematic because there is no clear mechanism of identifying and further updating the merged branches. This could be an interesting future work to explore.

%\noindent This \emph{multiple overriding} non-othogonal extension shows that our model is friendly to different kinds of language extensions.

Other typical non-orthogonal extensions to \MIM{} could be to have fields.
The design of \MIM{} can be viewed as a variant of Java 8 with default methods which allows for unintentional method conflicts.
Like Java interfaces and traits, state is forbidden in \MIM{}. There are some inheritance models that also account for fields, such as C++ that uses virtual inheritance~\cite{ellis1990annotated}. In our model, however, we can perhaps borrow the idea of \textit{interface-based programming}~\cite{classless}, which models state with abstract state operations. This can be realized by extending our current model with static methods and anonymous classes from Java. However such an extension requires more thought, so we leave it to future work.

\subsection{Loosening the Model: Reject Early or Reject Later?}\label{subsec:loosen}

\MIM{} rejects the following case of diamond inheritance:
\begin{lstlisting}
interface A { void m() {...} }
interface B extends A { void m() {...} }
interface C extends A { void m() {...} }
interface D extends B, C {}
\end{lstlisting}
Here both \lstinline|B.m| and \lstinline|C.m| override \lstinline|A.m|, and \lstinline|D| inherits
both conflicting methods without an explicit override. In this case, automatically merging 
the two methods (to achieve diamond inheritance) is not possible, which is why many 
models (like traits and Java 8) reject such programs. Moreover, keeping the two 
method implementations in \lstinline|D| is problematic. 
In essence, hierarchical information is 
not helpful to disambiguate later method calls, since the two methods share 
the same origin (\lstinline|A.m|). 
Our calculus rejects such conflicts by the \textsc{(T-Intf)} rule, where \lstinline|D| is considered to be ill-formed. 
We believe that rejecting \lstinline|D| follows the principle of models like traits and Java 8 interfaces, 
where the language/type-system is meant to alert the programmer for a possible conflict early.

Nonetheless,
 C++ accepts the definition of \lstinline|D|, but forbids later upcasts from \lstinline|D| to \lstinline|A| because of ambiguity. Our language is more conservative on definitions of interfaces compared to C++, but on the upside, upcasts are not rejected. We could also loosen the model to accept definitions such as \lstinline|D|, and perform ambiguity check on upcasts and other expressions. Then, we would need to handle more cases than C++ because of the complication caused by the hierarchical overriding feature. 
% For instance, in the above program assume that $B.m$ and $C.m$ are hierarchical overrides on $A$. Then such a case is still a diamond.

\begin{comment}
Another case is about the restriction on hierarchical overrides. In our language model there are two types of methods (though they share the same syntax): original methods and hierarchical overrides. We require that a hierarchical override should only override an original method. Yet from the idea of encapsulation and modularity on programs, we could also allow an override to work on another override, which forms a series of overrides.
\end{comment}

%\haoyuan{more. 1) Path Invocation, more generalized diamond inheritance: not clear; other method resolution techniques: no example; better code reuse with other kinds of method invocation: no example. }